### 2023-03-02 chapters/conclusion.tex
outline in imli

\begin{itemize}
	\item 
	\textbf{Post-hoc Explainability.} Black-box classifiers are hard to understand and trust, especially in safety-critical domains. In this thesis, we have designed interpretable classifiers that are explainable by design. We can use these classifiers, such as {\imli} and {\crr}, to generate post-hoc explanations for black-box classifiers. In this context, logic-based explanations, such as abduction-based explanations~\cite{ignatiev2019abduction}, Anchor~\cite{ribeiro2018anchors}, and LORE~\cite{guidotti2018local} have demonstrated to be effective for this purpose. Therefore, we plan to test how well {\imli} can synthesize explanations with large samples and how well {\crr} can generate more expressive explanations.

	\item \textbf{Fairness Auditing.} Machine learning has demonstrated significant success as a predictive system in automated decision-making. However, quantifying and mitigating bias in these systems is still challenging, e.g. bias in generative machine learning and large language models~\cite{abid2021persistent,nadeem2020stereoset,vig2020investigating}. As a result, there has been significant research interest in auditing classifiers for bias, from standard supervised learning to deep neural networks, computer vision, large language models and so on. Thus, we need a fairness auditing framework~\cite{ruf2021towards,yan2022active} with formal guarantees. One task of this framework is to categorize different definitions of fairness metrics in the literature. This would help stakeholders pick the right definition of fairness for their application~\cite{ruf2021towards}. We plan to extend formal fairness verification beyond group fairness, such as individual fairness~\cite{john2020verifying}, causal fairness~\cite{pan2021explaining,zhang2018fairness}, and counterfactual fairness~\cite{wu2019counterfactual,chiappa2019path}. We also plan to cover more classes of machine learning classifiers that are hard or impossible to encode as CNF, such as deep neural networks. Additionally, we aim to extend fairness verification beyond tabular data to more complex data such as images and texts.

>>>>

\begin{itemize}
	\item 
	\textbf{Post-hoc Explainability.} Black-box classifiers are hard to understand and trust, especially in safety-critical domains. In this thesis, we have designed interpretable classifiers that are explainable by design. We can use these classifiers, such as {\imli} and {\crr}, to generate post-hoc explanations for black-box classifiers. In this context, logic-based explanations, such as abduction-based explanations~\cite{ignatiev2019abduction}, Anchor~\cite{ribeiro2018anchors}, and LORE~\cite{guidotti2018local} have demonstrated to be effective in generating post-hoc explanations of black-box classifiers. Therefore, we plan to test how well {\imli} can synthesize explanations with large samples and how well {\crr} can generate more expressive explanations.

	\item \textbf{Fairness Auditing.} Machine learning has demonstrated significant success as a predictive system in automated decision-making. However, quantifying and mitigating bias in these systems is still challenging, e.g. bias in generative machine learning and large language models~\cite{abid2021persistent,nadeem2020stereoset,vig2020investigating}. As a result, there has been significant research interest in auditing classifiers for bias, from standard supervised learning to deep neural networks, computer vision, large language models and so on. Thus, we need a fairness auditing framework~\cite{ruf2021towards,yan2022active} with formal guarantees. One task of this framework is to categorize different definitions of fairness metrics in the literature. This would help stakeholders pick the right definition of fairness for their application~\cite{ruf2021towards}. In addition, we aim to extend formal fairness verification beyond group fairness, such as individual fairness~\cite{john2020verifying}, causal fairness~\cite{pan2021explaining,zhang2018fairness}, and counterfactual fairness~\cite{wu2019counterfactual,chiappa2019path}. We also plan to cover more classes of machine learning classifiers that are hard or impossible to encode as CNF, such as deep neural networks. Additionally, we aim to extend fairness verification beyond tabular data to more complex data such as images and texts.

### 2023-03-02 chapters/interpretability/imli/introduction.tex
outline in imli

%We organize the rest of the chapter as follows. We discuss related literature in Section~\ref{interpretability_imli_sec:related} and preliminaries in Section~\ref{interpretability_imli_sec:preliminaries}. In Section~\ref{interpretability_imli_sec:problem}, we formally define our learning problem and provide a MaxSAT-based solution in Section~\ref{interpretability_imli_sec:baseline}. We improve the MaxSAT-based formulation through an incremental learning technique in Section~\ref{interpretability_imli_sec:incremental_learning}. In Section~\ref{interpretability_imli_sec:application}, we apply {\imli} in learning different interpretable classifiers. We then discuss our experimental results in Section~\ref{interpretability_imli_sec:experiments} and conclude in Section~\ref{interpretability_imli_sec:conclusion}.

>>>>

We organize the rest of the chapter as follows. We discuss related literature in Chapter~\ref{interpretability_imli_sec:related}. In Chapter~\ref{interpretability_imli_sec:problem}, we formally define interpretable rule-based classification problem and discuss a MaxSAT-based solution proposed by~\cite{malioutov2018mlic} in Chapter~\ref{interpretability_imli_sec:baseline}. In Chapter~\ref{interpretability_imli_sec:incremental_learning}, we discuss the key technical contribution of this chapter, {\imli}, an improved MaxSAT-based learning formulation based on incremental solving. In Chapter~\ref{interpretability_imli_sec:application}, we apply {\imli} in learning different interpretable classifiers. We then discuss our experimental results in Chapter~\ref{interpretability_imli_sec:experiments} and conclude in Chapter~\ref{interpretability_imli_sec:conclusion}.
